On the Functionality of the Multilayer Perceptron

### 2020-11-30 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
Added new figures, Updates to Thesis

\begin{algorithm}[H]
\caption{Backwards propagation system, in a standard densely connected deep neural network. Each iteration in the \textit{for-loop} computes the gradient of the cost function $J$ with respect to the weight and bias arrays. Each element in those arrays $dW$ and $db$ is the discrete gradient of the cost due to that parameter. A practical application of this algorithm should include batches of samples instead of a single sample and a regularizing function at each step.}
\label{algBackProp}

@@

\FOR {$L-1,L-2,...,2,1$}
	\item Compute gradient w.r.t pre-nonlinear activation portion of layer function
	\item $dx^{(l)} \leftarrow \nabla_{Z^{(l)}}J = dx^{(l)} \odot \partial\sigma^{(l)}[ Z^{(l)} ]$
	\item Compute gradient w.r.t weighting and bias elements
	\item $db \leftarrow \nabla_{b^{(l)}}J = dx^{(l)}$
	\item $dW \leftarrow \nabla_{W^{(l)}}J = dx^{(l)} \cdot X^{(l-1)}$

@@

\begin{algorithmic}
\REQUIRE Step size $\alpha$
\REQUIRE Small constant $\delta$ for numerical stabilization, usually about $10^{-7}$.
\REQUIRE constants $\rho_1$, $\rho_2$ used got exponential decay rates, usually $0.9$ and $0.999$ respectively.
\REQUIRE Subroutine/function to compute gradient of cost function.
\REQUIRE Mini-batch size, $m$
\REQUIRE Stopping criterion $S$

@@

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-parameter of $b = 64$ samples per batch, and have developed $p = 24$ unique classification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 20)}
\end{equation}
This 2D array is referred to as a \textit{design matrix} \cite{James,Loy}. Any other shape will be rejected by the model, and an error is raised.

@@

\end{tabular}
\end{center}
\caption{Instruments categories used in the classification task. Note that the network uses only integers given by "Class index" to indentify sources. The string "Class Name" is keep for human readability}
\end{figure}

@@

\paragraph*{}The zero crossing rate (ZXR) of a signal or frame is use to measure how many times that a signal crosses it's equilibrium point. This can be computed per total sound wave, per analysis-frame, or per unit time. This feature is most commonly associated with differentiating speech from music, because speech presents a more jagged and often less periodic waveform \cite{Kahn,Liu,Zhang} that musical instruments do.

@@

\paragraph*{}Rather than producing filter banks based on the linear Hertz scale, the frequency axis of the signal is transformed into units of \textit{Mels}, which is used to account for the non-linearity in human pitch perception \cite{Serizel,Kahn}. Filter banks are produced to be evenly spaced on the Mel scale, and then transformed back into the Hertz scale. This has the effect of producing triangular filter banks with grow in width as the frequency increases. The Hertz to Mel and Mel to Hertz transforms are given \cite{Serizel,Kahn}:
\begin{equation}
\label{eqn-HztoMel}
M_f[h] = 2595 \log_{10}\big(1+ \frac{h}{700}\big)
\end{equation}
\begin{equation}
\label{eqn-MeltoHz}
H_f[m] = 700 \big(10^{(\frac{m}{2595})}-1\big)
\end{equation}
Where $M_f$ is the frequency in units of Mels, given $[h]$, a frequency in Hertz, and $H_f$ is the frequency in Hertz given $[m]$ a frequency in Mels.

@@

\paragraph*{}However, it is more useful to weight the confusion matrix according to class occurrence. For example, if we had a $2 \times 2$ matrix with entries $100$ and $20$ in the main diagonal, it seems that class $A$ has a much higher prediction success than class $B$. However, if there are $200$ elements in class $A$, and only $21$ in class $B$, then we see that class $A$ has a $50\%$ classification rate, while class $B$ has a $95\%$ classification rate, To Combat this, we typically normalize each confusion matrix by the number of samples in each "actual" class. Put simply, we all elements in a given row by the sum of that row.

@@

\subsubsection{Micro F1-Score}

@@

\subsubsection{Accuracy Score \&}

@@

\paragraph*{}For a multi-class problem, we present accuracy over \textit{all} classes simultaneously. It is the ratio of the sum of the main diagonal to the sum or all entries in the matrix:
\begin{equation}
\label{eqn-AvgAccuracy}
\text{Accuracy} = \frac{\sum_{i}C_{i,i}}{\sum_{i}\sum_{j}C_{i,j}}
\end{equation}
Where all sum indexes goes from $0$ to $k-1$.

@@

\begin{figure}[h]
\begin{center}
%\includegraphics[scale=0.2]{•}
\textcolor{red}{Precision Score Goes Here}
\textcolor{red}{Recall Score Goes Here}
\end{center}
\caption{The precision score (a) and recall score (b) increases with each training step}
\label{fig-PrecisionRecallScores}
\end{figure}

@@

\paragraph*{}To formally produce classification predictions on the chaotic synthesizer wave forms, we must use all of the principles outlined in the previous sections. Using the described features in sec. (\ref{sec-Features}) and appropriate architecture sec. (\ref{subsec-Architecture}), We run a

we produce a neural network \textit{ChaoticSynthClassifier} and train it on the roughly $18,000$ audio files that serve as the training set.

>>>>

\begin{algorithm}[H]
\caption{Backwards propagation system, in a standard densely connected deep neural network. Each iteration in the \textit{for-loop} computes the gradient of the cost function $J$ with respect to the weight and bias arrays in a given layer $(l)$. Each element in those arrays $dW$ and $db$ is the discrete gradient of the cost due to that parameter. A practical application of this algorithm should include batches of samples instead of a single sample and a regularizing function at each step.}
\label{alg-BackProp}

@@

\FOR {$l \in [L-1,L-2,...,2,1]$}
	\item Compute gradient w.r.t pre-nonlinear activation portion of layer function
	\item $dx^{(l)} \leftarrow \nabla_{Z^{(l)}}J = dx^{(l)} \odot \partial\sigma^{(l)}[ Z^{(l)} ]$
	\item Compute gradient w.r.t weighting and bias elements
	\item $db \leftarrow \nabla_{b^{(l)}}J = dx^{(l)}$
	\item $dW \leftarrow \nabla_{W^{(l)}}J = dx^{(l)} \cdot X^{(l-1)}$

@@

\begin{algorithmic}
\REQUIRE Step size $\alpha$
\REQUIRE Small constant $\delta$ for numerical stabilization, usually about $10^{-7}$.
\REQUIRE Constants $\rho_1$, $\rho_2$ used track exponential decay rates, usually $0.9$ and $0.999$ respectively.
\REQUIRE Subroutine/function to compute gradient of cost function See Alg. (\ref{alg-BackProp})
\REQUIRE Mini-batch size, $m$
\REQUIRE Stopping criterion $S$

@@

\paragraph*{}The input layer of the perceptron accepts a 2-Dimensional array. The axes, in order of indexing, represent (i) the size of the \textit{mini-batch} of samples, (ii) the number of features for each sample. We use the same model hyper-parameter of $b = 64$ samples per batch, and have developed $p = 24$ unique classification features that have been derived from time-space and frequency-space representations of the audio file data. We can denote the 2D shape if the input object into this branch as:
\begin{equation}
\label{eqn-shapeX2}
X_2 \in \mathbb{R}^{(64 \times 24)}
\end{equation}
This 2D array is referred to as a \textit{design matrix} \cite{James,Loy}. Any other shape will be rejected by the model, and an error is raised.

@@

\end{tabular}
\end{center}
\caption{Instruments categories used in the classification task. Note that the network uses only integers given by "Class index" to indentify sources. The string "Class Name" is keep for human readability}
\label{fig-ClassList}
\end{figure}

@@

\paragraph*{}The zero crossing rate (ZXR) of a signal or frame is use to measure how many times that a signal crosses it's equilibrium point. This can be computed per total sound wave, per analysis-frame, or per unit time. This feature is most commonly associated with differentiating speech from music, because speech presents a more jagged and often less periodic waveform that musical instruments do \cite{Khan,Liu,Zhang} .

@@

\paragraph*{}Rather than producing filter banks based on the linear Hertz scale, the frequency axis of the signal is transformed into units of \textit{Mels}, which is used to account for the non-linearity in human pitch perception \cite{Serizel,Khan}. Filter banks are produced to be evenly spaced on the Mel scale, and then transformed back into the Hertz scale. This has the effect of producing triangular filter banks with grow in width as the frequency increases. The Hertz to Mel and Mel to Hertz transforms are given \cite{Serizel,Kahn}:
\begin{equation}
\label{eqn-HztoMel}
M_f[h] = 2595 \log_{10}\big(1+ \frac{h}{700}\big)
\end{equation}
\begin{equation}
\label{eqn-MeltoHz}
H_f[m] = 700 \big(10^{(\frac{m}{2595})}-1\big)
\end{equation}
Where $M_f$ is the frequency in units of Mels, given $[h]$, a frequency in Hertz, and $H_f$ is the frequency in Hertz given $[m]$ a frequency in Mels.

@@

\paragraph*{}However, it is more useful to weight the confusion matrix according to class occurrence. For example, if we had a $2 \times 2$ matrix with entries $100$ and $20$ in the main diagonal, it seems that class $A$ has a much higher prediction success than class $B$. However, if there are $200$ elements in class $A$, and only $21$ in class $B$, then we see that class $A$ has a $50\%$ classification rate, while class $B$ has a $95\%$ classification rate. Consider the list of instrument classes in Fig.(\ref{fig-ClassList}) and notice how all instruments are not equally represented. This creates a training bias such that the model is more heavily trained on a particular class compared to others. This may also show when making predictions, where the classifier is more likely to predict samples that it has been trained on more \cite{James,Loy}. To Combat this, we typically normalize each confusion matrix by the number of samples in each "actual" class. Put simply, we divide each element in a given row by the \textit{sum} of that row.

@@

\subsubsection{F1-Score}

\paragraph*{}F1-score (also called F-score) is the harmonic mean of precision and recall scores and it also bounded on $[0,1]$ with a higher score being favorable \cite{Geron}. Often, the two metrics can be through of as somewhat \textit{exclusive} to each other - the higher one, the lower the other. Compare this with the idea that the more \textit{sensitive} the model is, the less \textit{specific} it is, and vice-versa. This phenomenon is called the \textit{precision-recall tradeoff} \cite{Geron,James}. We compute the $F1$ score for a class $k$ as:
\begin{equation}
\label{eqn-F1Score}
\text{F1}_{j} = 2 \times \frac{\text{Prec}_j + \text{Recall}_j}{\text{Prec}_j \times \text{Recall}_j}
\end{equation}

\paragraph*{}The $F1$ favors models with both a high precision and high recall score. Some models allow for the ability to adjust the threshold of the decision function. This means that by changing a few parameters that are external to the classifier itself, we can change how sensitive or specific a model is when making it's class decision. In the case of this multi-category classifier, this is somewhat like modifying the generated decision boundaries as to change the outputted probability distribution. In case cases, it is useful to use the $F1$ score to find the set of hyper-parameters that allow for the classifier to produce both the highest possible precision \textit{and} recall score.

@@

\subsubsection{Accuracy Score}

@@

\paragraph*{}For a multi-class problem, present accuracy over each individual class, called \textit{micro-accuracy} \cite{James}. The accuracy for a class $j$ is the correct predictions related to the class divided by the total predictions and total counts of the class. We express this mathematically by using sums over each row, each column, and the main diagonal as such:
\begin{equation}
\label{eqn-Accuracy}
\text{Accuracy}_j = \frac{\sum_{i}C_{i,i}}{\sum_{i}C_{i,i} + \sum_{i \neq j}C_{i,j} + \sum_{i \neq j}C_{j,i}}
\end{equation}
Where all sum indexes goes from $0$ to $k-1$.

\paragraph*{}We can also present \textit{global accuracy} or \textit{macro accuracy} by computing the acurracy over the full confusion matrix. This would then by the ratio of all correct predictions (the main diagonal) to all predictions (the sum of the matrix). We can express this as:
\begin{equation}
\text{Accuracy} = \frac{\sum_{i}C_{i,i}}{\sum_{i}\sum_{j}C_{i,j}}
\end{equation}

@@

\begin{figure}[h]
\begin{center}

\end{center}
\caption{The precision score (a) and recall score (b) increases with each training step}
\label{fig-PrecisionRecallScores}
\end{figure}

@@

\subsection{Executing Cross Validation}

\paragraph*{}To formally produce classification predictions on the chaotic synthesizer wave forms, we must use all of the principles outlined in the previous sections. Using the described features in sec. (\ref{sec-Features}) and appropriate architecture sec. (\ref{subsec-Architecture}), We run a $K=10$ folds cross validation program as in Alg.(\ref{alg-CrossValidation}). For each of the $10$ models, we produce a standard confusion matrix, an compute the (i) accuracy score, (ii) precision score, (iii) recall score, and (iv) F1 score. For each of the $37$ classes, we compute the arithmetic mean of the metrics across the $10$ folds. We show the results of this in Figs. (\ref{fig-HighWindsScores}), (\ref{fig-LowWindsScores}), (\ref{fig-BrassScores})

\paragraph*{}Additionally, we report the average of each four metrics across all classes for each of the $10$ splits in Fig. (\ref{fig-ValidationTable}). We also present a more condensed version of the data by showing the average, minimum, maximum, and variance of each score in Fig. (\ref{fig-ValidationStats}). Each value is on the interval $[0,1]$, with higher values being more favorable.
\begin{figure}[H]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Split Index} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
$0$ & 0.992969341 & 0.849762613	& 0.853483643 & 0.838007935 \\ \hline
$1$ & 0.993086645 & 0.857063114	& 0.865737308 & 0.854069688 \\ \hline
$2$ & 0.994175444 & 0.858936065	& 0.854231756 & 0.851319866 \\ \hline
$3$ & 0.993883291 & 0.887070101	& 0.901606142 & 0.882074022 \\ \hline
$4$ & 0.992941956 & 0.884115922	& 0.902609597 & 0.884775112 \\ \hline
$5$ & 0.992941956 & 0.884115922	& 0.902609597 & 0.884775112 \\ \hline
$6$ & 0.994315799 & 0.856425689	& 0.842680234 & 0.834139910 \\ \hline
$7$ & 0.993057475 & 0.869049669	& 0.867171387 & 0.856152544 \\ \hline
$8$ & 0.992979711 & 0.843322939	& 0.866599329 & 0.846637817 \\ \hline
$9$ & 0.992799981 & 0.842635691	& 0.877821096 & 0.832496007 \\ \hline
\end{tabular}
\end{center}
\caption{Metrics averaged over all classes for $K=10$ validation splits}
\label{fig-ValidationTable}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Measure} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
\textbf{Average} & 0.993385794 & 0.860133646 & 0.872055273 & 0.853799365 \\ \hline
\textbf{Minimum} & 0.992799981 & 0.842635691 & 0.842680234 & 0.832496007 \\ \hline
\textbf{Maximum} & 0.994315799 & 0.887070101 & 0.902609597 & 0.884775112 \\ \hline
\textbf{Variance}& 2.88066E-07 & 0.000215270 & 0.000373321 & 0.000293214 \\ \hline
\end{tabular}
\end{center}
\caption{Metric statisitcs over all classes  and all validation splits}
\label{fig-ValidationStats}
\end{figure}

	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/AltoFlute}
	\caption{Alto Flute}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/AltoSax}
	\caption{Alto Saxophone}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/BbClarinet}
	\caption{$B\flat$ Clarinet}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/EbClarinet}
	\caption{$E\flat$ Clarinet}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/EnglishHorn}
	\caption{English Horn}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Flute}
	\caption{Flute}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Oboe}
	\caption{Oboe}
	\end{subfigure}		
	\begin{subfigure}[b]{0.45 \textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/SopSax}
	\caption{Soprano Saxophone}
	\end{subfigure}

\caption{Performance Metrics Across all High Ranged Woodwind Instruments}
\label{fig-HighWindsScores}
\end{figure}

\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/BassClarinet}
	\caption{Bass Clarinet}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/BassFlute}
	\caption{Bass Flute}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Bassoon}
	\caption{Bassoon}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/ContraBassoon}
	\caption{Contra Bassoon}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Saxophone}
	\caption{Tenor Saxophone}
	\end{subfigure}
\caption{Performance Metrics Across all Mid and Low Ranged Woodwind Instruments}
\label{fig-LowWindsScores}
\end{figure}

\begin{figure}[H]
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/BassTrombone}
	\caption{Bass Trombone}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Horn}
	\caption{French Horn}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/TenorTrombone}
	\caption{Tenor Trombone}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Trombone}
	\caption{Trombone}
	\end{subfigure}	
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Trumpet}
	\caption{$B\flat$ Trumpet}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{../FiguresClasses/Tuba}
	\caption{Tuba}
	\end{subfigure}
\caption{Performance Metrics Across all Brass Instruments}
\label{fig-BrassScores}
\end{figure}

\paragraph*{}With cross validation completed and the results

\subsection{•}
